---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "London Weather - How bad is it really?"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Ruan Geldenhuys"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Stellenbosch, South Africa" # First Author's Affiliation
Email1: "22550801\\@sun.ac.za" # First Author's Email address



# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
 In order to convinve my friend not to move to the UK, I did some data analysis on London's weather patterns. Since she is not a big reader, I will keep my discussion to a minimum.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(viridis)
library(xtable)

# Notice that as you are working in a .Rproj file (I am assuming you are) - the relative paths of your directories start at your specified root.
# This means that when working in a .Rproj file, you never need to use getwd() - it is assumed as your base root automatically.


```

```{r}
londonWeather <- read.csv('data/London/london_weather.csv')
UKmonthlyDF <- read.csv('data/London/UKMonthly_Detailed.csv')
londonDF <- londonWeather %>%
    mutate(date = ymd(date)) %>%
    mutate(date = format(date, "%Y/%m/%d")) %>%
    mutate(month = month(date, label = TRUE))
source('code/PlotTempPrecip.R')
source('code/coveragePlot.R')
source('code/rainyDays.R')
source('code/rainComparePlot.R')

```

# Temperature and Precipation

The graph below shows a breakdown of temperature and rain in London throughout the year.

```{r  fig.width=7, fig.height=5}
plotTempPrecip(londonDF)
```

It's clear that London is a very cold and wet place with the maximum average temperature being 23.7 and having dearly 2 inches (50mm) of rain all year round.

# Sunshine, or rather the lack thereof

The plot below shows the average cloud coverage in London by month.

```{r}
coveragePlot(londonDF, "Monthly Cloud Coverage - London")
```

Sunny days are few in London, as most days are partly cloudy or overcast, as can be seen in the graph above.

# London vs Cape Town

Using data from the Southern Africa Tourism Services Association we can compare London and Cape Town.

```{r, results='asis'}
rainTable <- options(xtable.comment = FALSE)
rainTable <- rainyTable(londonDF)

print(rainTable, caption.placement = 'top')
```
```{r}
rainComparePlot(londonDF)
```

It is clear from the table and the graph that London experiences vastly more rain than Cape Town. London in its driest month gets more rain than Cape Town in its wettest month.
